---
title: "Confirmed Cases"
---

<style type="text/css">
div.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r initialize, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
#Install/Activate necessary packages
source("scripts/00M_Fonts_And_Colors.R")
source("scripts/01M_Function_Declarations.R")
source("scripts/02M_Data_Prep.R")

# Run all scripts relevant to the confirmed cases
# of COVID-19 needed for this section
source("scripts/cases/03C_Daily.R")
source("scripts/cases/04C_Symptomatic.R")
source("scripts/cases/05C_Asymptomatic.R")
source("scripts/cases/06C_Age_Group.R")
source("scripts/cases/07C_Avg_Age.R")
source("scripts/cases/08C_Gender.R")
source("scripts/cases/09C_Race.R")
source("scripts/cases/10C_Ethnicity.R")
source("scripts/cases/11C_Zip_Codes.R")

ttl = list() # List of totals
ftnotes = list() # List of footnotes
pct = list() # List of percentages
```

```{r cases footnotes, message=FALSE, warning=FALSE, include=FALSE}
# Total of all cases in each category (where known)
ttl <- with(
  processed_data$cases,
  list(
    cases = list(
      cases   = sum(daily$Cases),
      symp    = sum(symp$Cases),
      asymp   = sum(asymp$Cases),
      age     = sum(age_gr$Cases),
      gender  = sum(gender$Cases),
      race    = sum(race$Cases),
      ethn    = sum(ethn$Cases),
      top_zip = sum(top_zip$Cases)
    )
  )
)

# Calculate the percentage values for age groups and race
pct$cases <- c(
  with(
  processed_data$cases$age_gr,
  list(
    age0 = Cases[Age.Group == "0-17"]  / ttl$cases$age,
    age1 = Cases[Age.Group == "18-44"] / ttl$cases$age,
    age2 = Cases[Age.Group == "45-64"] / ttl$cases$age,
    age3 = Cases[Age.Group == "65-74"] / ttl$cases$age,
    age4 = Cases[Age.Group == "75+"]   / ttl$cases$age
  )
),
  with(
    processed_data$cases$race,
    list(
      white  = Cases[Race == data_validation$race$White]  / ttl$cases$race,
      black  = Cases[Race == data_validation$race$Black]  / ttl$cases$race,
      asian  = Cases[Race == data_validation$race$Asian]  / ttl$cases$race,
      indian = Cases[Race == data_validation$race$Native] / ttl$cases$race,
      other  = Cases[Race == data_validation$race$Other]  / ttl$cases$race
    )
  )
)

ftnotes$cases <- c(
  with(
    pct$cases,
    # Format the percentages to one decimal point and add
    # a % symbol after the value.
    list(
      age0   = sprintf("%0.1f%%", age0 * 100),
      age1   = sprintf("%0.1f%%", age1 * 100),
      age2   = sprintf("%0.1f%%", age2 * 100),
      age3   = sprintf("%0.1f%%", age3 * 100),
      age4   = sprintf("%0.1f%%", age4 * 100),
      white  = sprintf("%0.1f%%", white  * 100),
      black  = sprintf("%0.1f%%", black  * 100),
      asian  = sprintf("%0.1f%%", asian  * 100),
      indian = sprintf("%0.1f%%", indian * 100),
      other  = sprintf("%0.1f%%", other  * 100)
    )
  ),
  with(
    ttl$cases,
    # Format all values that use values from ttl$cases
    list(
      unk_age     = format(as.numeric(cases - age), big.mark = ","),
      unk_gender  = format(as.numeric(cases - gender), big.mark = ","),
      unk_race    = format(as.numeric(cases - race), big.mark = ","),
      unk_ethn    = format(as.numeric(cases - ethn), big.mark = ","),
      top_zip     = format(as.numeric(top_zip), big.mark = ","),
      pct_top_zip = sprintf("%0.1f%%", top_zip / cases * 100),
      asymp       = format(as.numeric(asymp), big.mark = ","),
      unk_symp    = format(as.numeric(cases - symp - asymp), big.mark = ","),
      symp        = format(as.numeric(symp), big.mark = ",")
    )
  ),
  avg_age = sprintf("%0.1f", mean(processed_data$cases$avg_age$Age))
)
```

## Confirmed Cases of COVID-19 {.tabset}

### Daily Cases
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$daily
```
**Note: The dates in 2020 of April 24th, April 29th, December 17th, and in 2021 of March 17th and December 29th were dates when laboratories uploaded backlogged data.**

### Symptomatic Cases
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$symp
```
**Note: This graph does not reflect the most recent cases due to case investigation and data processing. Individuals who were asymptomatic (`r ftnotes$cases$asymp`), in data processing, were lost to follow up and/or had an unknown date of first symptom (`r ftnotes$cases$unk_symp`) or reported a symptom onset date prior to December 1, 2019 (<5) were excluded from analysis. Cases represented in graph (n=`r ftnotes$cases$symp`).**

### Asymptomatic Cases
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$asymp
```
**Note: This graph does not reflect the most recent cases due to case investigation and data processing. Represented in graph (n=`r ftnotes$cases$asymp`) reporting no symptoms at the time of interview.**

## Demographics {.tabset}

### Age
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$age_gr
```
**Note: Ages were calculated using laboratory reported or case reported date of birth. Missing data (n=`r ftnotes$cases$unk_age`). The percent per age group is: 0-17 = `r ftnotes$cases$age0`; 18-44 = `r ftnotes$cases$age1`; 45-64 = `r ftnotes$cases$age2`; 65-74 = `r ftnotes$cases$age3`; and 75+ = `r ftnotes$cases$age4`.**

<div align = "center">
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE}
plots$cases$avg_age
```
</div>
**Note: Ages were calculated using laboratory reported or case reported date of birth. Missing data (n=`r ftnotes$cases$unk_age`). The calculated mean of known ages of confirmed cases of COVID-19 is `r ftnotes$cases$avg_age`.**

### Gender
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$gender
```
**Note: Gender is self-reported field. Missing or Unknown when laboratory reporting did not provide this information and case has not been reached yet (n=`r ftnotes$cases$unk_gender`).**

### Race
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$race
```
**Note: Race is largely self-reported data and are identified through case investigation. Data are incomplete due to delays in reporting, data entry, unknown or undetermined risk factors and persons lost to follow up (n=`r ftnotes$cases$unk_race`).<br>The percent per group is: White = `r ftnotes$cases$white`; Other Race = `r ftnotes$cases$other`; Black = `r ftnotes$cases$black`; Asian or Pacific Islander or Native Hawaiian = `r ftnotes$cases$asian`; and Native American or Alaskan Native = `r ftnotes$cases$indian`.**

### Ethnicity
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$ethn
```
**Note: Ethnicity is largely self-reported data and are identified through case investigation. Data are incomplete due to delays in reporting, data entry,unknown, or undetermined risk factors and persons lost to follow-up (n=`r ftnotes$cases$unk_ethn`).**

## Municipalities
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$top_zip
```
**Note: Data is incomplete due to ongoing case investigation. Twenty most reported zip codes represented in graph (n=`r ftnotes$cases$top_zip`) representing `r ftnotes$cases$pct_top_zip` of cases.**

