---
title: "COVID-19 Dashboard"
author: Orange County Dept of Health
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smoot_scroll: true
runtime: shiny
---

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'text-indent': num_ul * 10, 'padding-left': 0});
  });
});
</script>

<style type="text/css">
div.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r initialize, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = '..')
```

```{r setup, include=FALSE}
#Install/Activate necessary packages
source("scripts/00M_Fonts_And_Colors.R")
source("scripts/01M_Function_Declarations.R")
source("scripts/02M_Data_Prep.R")

# Run all scripts relevant to the confirmed cases
# of COVID-19 needed for this section
source("scripts/cases/03C_Daily.R")
source("scripts/cases/04C_Symptomatic.R")
source("scripts/cases/05C_Asymptomatic.R")
source("scripts/cases/06C_Age_Group.R")
source("scripts/cases/07C_Avg_Age.R")
source("scripts/cases/08C_Gender.R")
source("scripts/cases/09C_Race.R")
source("scripts/cases/10C_Ethnicity.R")
source("scripts/cases/11C_Zip_Codes.R")

# Run all scripts relevant to the deaths
# of COVID-19 needed for this section
source("scripts/deaths/03D_Date_of_Death.R")
source("scripts/deaths/04D_Gender.R")
source("scripts/deaths/05D_Race.R")
source("scripts/deaths/06D_Ethnicity.R")
source("scripts/deaths/07D_Age_Group.R")
source("scripts/deaths/08D_Avg_Age.R")
source("scripts/deaths/09D_Vaccination.R")
source("scripts/deaths/10D_Top_Zip.R")

ttl = list() # List of totals
ftnotes = list() # List of footnotes
pct = list() # List of percentages
```

## Confirmed Cases
```{r cases footnotes, message=FALSE, warning=FALSE, include=FALSE}
# Total of all cases in each category (where known)
ttl <- with(
  processed_data$cases,
  list(
    cases = list(
      cases   = sum(daily$Cases),
      symp    = sum(symp$Cases),
      asymp   = sum(asymp$Cases),
      age     = sum(age_gr$Cases),
      gender  = sum(gender$Cases),
      race    = sum(race$Cases),
      ethn    = sum(ethn$Cases),
      top_zip = sum(top_zip$Cases)
    )
  )
)

# Calculate the percentage values for age groups and race
pct$cases <- c(
  with(
  processed_data$cases$age_gr,
  list(
    age0 = Cases[Age.Group == "0-17"]  / ttl$cases$age,
    age1 = Cases[Age.Group == "18-44"] / ttl$cases$age,
    age2 = Cases[Age.Group == "45-64"] / ttl$cases$age,
    age3 = Cases[Age.Group == "65-74"] / ttl$cases$age,
    age4 = Cases[Age.Group == "75+"]   / ttl$cases$age
  )
),
  with(
    processed_data$cases$race,
    list(
      white  = Cases[Race == data_validation$race$White]  / ttl$cases$race,
      black  = Cases[Race == data_validation$race$Black]  / ttl$cases$race,
      asian  = Cases[Race == data_validation$race$Asian]  / ttl$cases$race,
      indian = Cases[Race == data_validation$race$Native] / ttl$cases$race,
      other  = Cases[Race == data_validation$race$Other]  / ttl$cases$race
    )
  )
)

ftnotes$cases <- c(
  with(
    pct$cases,
    # Format the percentages to one decimal point and add
    # a % symbol after the value.
    list(
      age0   = sprintf("%0.1f%%", age0 * 100),
      age1   = sprintf("%0.1f%%", age1 * 100),
      age2   = sprintf("%0.1f%%", age2 * 100),
      age3   = sprintf("%0.1f%%", age3 * 100),
      age4   = sprintf("%0.1f%%", age4 * 100),
      white  = sprintf("%0.1f%%", white  * 100),
      black  = sprintf("%0.1f%%", black  * 100),
      asian  = sprintf("%0.1f%%", asian  * 100),
      indian = sprintf("%0.1f%%", indian * 100),
      other  = sprintf("%0.1f%%", other  * 100)
    )
  ),
  with(
    ttl$cases,
    # Format all values that use values from ttl$cases
    list(
      unk_age     = format(as.numeric(cases - age), big.mark = ","),
      unk_gender  = format(as.numeric(cases - gender), big.mark = ","),
      unk_race    = format(as.numeric(cases - race), big.mark = ","),
      unk_ethn    = format(as.numeric(cases - ethn), big.mark = ","),
      top_zip     = format(as.numeric(top_zip), big.mark = ","),
      pct_top_zip = sprintf("%0.1f%%", top_zip / cases * 100),
      asymp       = format(as.numeric(asymp), big.mark = ","),
      unk_symp    = format(as.numeric(cases - symp - asymp), big.mark = ","),
      symp        = format(as.numeric(symp), big.mark = ",")
    )
  ),
  avg_age = sprintf("%0.1f", mean(processed_data$cases$avg_age$Age))
)
```

### Confirmed Cases of COVID-19 {.tabset}

#### Daily Cases
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$daily
```
<b>Note: The dates in 2020 of April 24th, April 29th, December 17th, and in 2021 of March 17th and December 29th were dates when laboratories uploaded backlogged data.</b><br><br>

#### Symptomatic Cases
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$symp
```
<b>Note: This graph does not reflect the most recent cases due to case investigation and data processing. Individuals who were asymptomatic (`r ftnotes$cases$asymp`), in data processing, were lost to follow up and/or had an unknown date of first symptom (`r ftnotes$cases$unk_symp`) or reported a symptom onset date prior to December 1, 2019 (<5) were excluded from analysis. Cases represented in graph (n=`r ftnotes$cases$symp`).</b><br><br>

#### Asymptomatic Cases
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$asymp
```
<b>Note: This graph does not reflect the most recent cases due to case investigation and data processing. Represented in graph (n=`r ftnotes$cases$asymp`) reporting no symptoms at the time of interview.</b><br><br>

### Demographics {.tabset}

#### Age
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$age_gr
```
<b>Note: Ages were calculated using laboratory reported or case reported date of birth. Missing data (n=`r ftnotes$cases$unk_age`).<br>
The percent per age group is: 0-17 = `r ftnotes$cases$age0`; 18-44 = `r ftnotes$cases$age1`; 45-64 = `r ftnotes$cases$age2`; 65-74 = `r ftnotes$cases$age3`; and 75+ = `r ftnotes$cases$age4`.</b><br>

<div align = "center">
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE}
plots$cases$avg_age
```
</div>
<b>Note: Ages were calculated using laboratory reported or case reported date of birth. Missing data (n=`r ftnotes$cases$unk_age`). The calculated mean of known ages of confirmed cases of COVID-19 is `r ftnotes$cases$avg_age`.</b><br><br>

#### Gender
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$gender
```
<b>Note: Gender is self-reported field. Missing or Unknown when laboratory reporting did not provide this information and case has not been reached yet (n=`r ftnotes$cases$unk_gender`).</b><br><br>

#### Race
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$race
```
<b>Note: Race is largely self-reported data and are identified through case investigation. Data are incomplete due to delays in reporting, data entry, unknown or undetermined risk factors and persons lost to follow up (n=`r ftnotes$cases$unk_race`).<br>The percent per group is: White = `r ftnotes$cases$white`; Other Race = `r ftnotes$cases$other`; Black = `r ftnotes$cases$black`; Asian or Pacific Islander or Native Hawaiian = `r ftnotes$cases$asian`; and Native American or Alaskan Native = `r ftnotes$cases$indian`.</b><br><br>

#### Ethnicity
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$ethn
```
<b>Note: Ethnicity is largely self-reported data and are identified through case investigation. Data are incomplete due to delays in reporting, data entry,unknown, or undetermined risk factors and persons lost to follow-up (n=`r ftnotes$cases$unk_ethn`).</b><br><br>

### Municipalities
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$cases$top_zip
```
<b>Note: Data is incomplete due to ongoing case investigation. Twenty most reported zip codes represented in graph (n=`r ftnotes$cases$top_zip`) representing `r ftnotes$cases$pct_top_zip` of cases.</b><br><br>

```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
# Complete the data set by adding zeros for any non-occurring
# data, places zero for count if a row with the date received
# and zip code does not exist for graphing purposes.
processed_data$cases$zip_daily <- raw_data$cases %>%
  filter(Zip.Code %in% data_validation$zip_codes$Zip.Code) %>%
  select(c(Date.Received, Zip.Code)) %>%
  cbind(z = 1) %>%
  tidyr::complete(Date.Received, Zip.Code, fill = list(z = 0)) %>%
  group_by(Date.Received, Zip.Code) %>%
  summarise(Cases = sum(z))

# Add in town/city names to each row
# Filters out any data that does not match a valid zip code
processed_data$cases$zip_daily <- merge(processed_data$cases$zip_daily, data_validation$zip_codes, by = 'Zip.Code')

# Dropdown list of selectable values to change the graph
selectInput(inputId = "id_town",
            label = "Zip Code",
            choices = sort(unique(processed_data$cases$zip_daily$Town.City)),
            selected = "Goshen"
)

plotlyOutput("town_city")

# Dynamically create the graph based on user input
output$town_city <- renderPlotly({
  # Create a temp variable to hold filtered data
  # and sort the data to be in date order
  temp_data <- filter(processed_data$cases$zip_daily, Town.City == input$id_town)
  temp_data <- temp_data[order(temp_data$Date.Received, decreasing = FALSE), ]
  
  # Create the chart title
  max_date <- format(as.Date(max(temp_data$Date.Received)), "%B %d, %Y")
  cases <- format(as.numeric(sum(temp_data$Cases)), big.mark = ",")
  
  chart_title <- str_glue(
    'Confirmed Cases of COVID-19
    {input$id_town}, NY as of {max_date} (n={cases})')
  
  y_max <- max(temp_data$Cases) * 1.05

  plots$zip_daily <- create_plotly_line(
    temp_data,
    temp_data$Date.Received,
    temp_data$Cases,
    my_colors$my_blue,
    "Date Received"
  )
  
  plots$zip_daily <- format_plotly(
    plots$zip_daily,
    chart_title,
    "Date Received",
    "Cases",
    "trace",
    y_max
  )
})
```
<b>Note: Data is incomplete due to ongoing case investigation.</b><br><br>

## Deaths
```{r death footnotes, message=FALSE, warning=FALSE, include=FALSE}
ttl$deaths <- with(
  processed_data$deaths,
  list(
    dod       = sum(dod$Deaths),
    gender    = sum(gender$Deaths),
    race      = sum(race$Deaths),
    ethn      = sum(ethn$Deaths),
    age_gr    = sum(age_gr$Deaths),
    vacc      = sum(vacc$Deaths),
    top_zip   = sum(top_zip$Deaths),
    after2021 = sum(subset(dod, Date.of.Death >= as.Date("2021-01-01"))$Deaths)
  )
)

pct$deaths <- c(
  # Calculate percentages for age group
  with(
    processed_data$deaths$age_gr,
    list(
      age0 = Deaths[Age.Group == "0-17"]  / ttl$deaths$age_gr,
      age1 = Deaths[Age.Group == "18-44"] / ttl$deaths$age_gr,
      age2 = Deaths[Age.Group == "45-64"] / ttl$deaths$age_gr,
      age3 = Deaths[Age.Group == "65-74"] / ttl$deaths$age_gr,
      age4 = Deaths[Age.Group == "75+"]   / ttl$deaths$age_gr
    )
  ),
  # Calculate percentages for race
  with(
    processed_data$deaths$race,
    list(
      white  = Deaths[Race == data_validation$race$White]  / ttl$deaths$race,
      black  = Deaths[Race == data_validation$race$Black]  / ttl$deaths$race,
      asian  = Deaths[Race == data_validation$race$Asian]  / ttl$deaths$race,
      indian = Deaths[Race == data_validation$race$Native] / ttl$deaths$race,
      other  = Deaths[Race == data_validation$race$Other]  / ttl$deaths$race
    )
  ),
  after2021 = ttl$deaths$after2021 / ttl$deaths$dod
)

# Format footnotes for the deaths section
ftnotes$deaths <- c(
  with(
    pct$deaths,
    # Format the percentages to one decimal point and add
    # a % symbol after the value.
    list(
      age0   = sprintf("%0.1f%%", age0 * 100),
      age1   = sprintf("%0.1f%%", age1 * 100),
      age2   = sprintf("%0.1f%%", age2 * 100),
      age3   = sprintf("%0.1f%%", age3 * 100),
      age4   = sprintf("%0.1f%%", age4 * 100),
      white  = sprintf("%0.1f%%", white  * 100),
      black  = sprintf("%0.1f%%", black  * 100),
      asian  = sprintf("%0.1f%%", asian  * 100),
      indian = sprintf("%0.1f%%", indian * 100),
      other  = sprintf("%0.1f%%", other  * 100),
      pct_after2021 = sprintf("%0.1f%%", after2021 * 100)
    )
  ),
  with(
    ttl$deaths,
    # Format all values that use values from ttl$deaths
    list(
      unk_race      = format(as.numeric(dod - race), big.mark = ","),
      unk_ethn      = format(as.numeric(dod - ethn), big.mark = ","),
      top_zip       = format(as.numeric(top_zip), big.mark = ","),
      pct_zip       = sprintf("%0.1f%%", top_zip / dod * 100),
      ttl_after2021 = format(as.numeric(after2021), big.mark = ",")
    )
  )
)
```

### Confirmed and Probable Deaths
```{r, out.width='100%', fig.height=5}
plots$deaths$dod
```
<b>Note: Due to delays in reporting recent data are incomplete.</b><br><br>

### Demographics {.tabset}

#### Age
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$deaths$age_gr
```
<b>Note: Due to delays in reporting recent data are incomplete. The percent per age group is: 0-17 = `r ftnotes$deaths$age0`; 18-44 = `r ftnotes$deaths$age1`; 45-64 = `r ftnotes$deaths$age2`; 65-74 = `r ftnotes$deaths$age3`; and 75+ = `r ftnotes$deaths$age4`.</b><br>

<div align = "center">
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE}
plots$deaths$avg_age

ftnotes$deaths$avg_age <- sprintf("%0.1f", mean(processed_data$deaths$avg_age$Age.At.Death))
```
</div>
<b>Note: Due to delays in reporting recent data are incomplete. Ages were calculated using laboratory reported or case reported date of birth. The calculated mean of known ages of confirmed cases of COVID-19 is `r ftnotes$deaths$avg_age`.</b><br><br>

#### Gender
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$deaths$gender
```
<b>Note: Due to delays in reporting recent data are incomplete.</b><br><br>

#### Race
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$deaths$race
```
<b>Note: Due to delays in reporting recent data are incomplete. Deaths with unknown race are not reported in the graph above (n=`r ftnotes$deaths$unk_race`). The percent per group is: White = `r ftnotes$deaths$white`; Black = `r ftnotes$deaths$black`; Other Race = `r ftnotes$deaths$other`; Asian or Pacific Islander or Native Hawaiian = `r ftnotes$deaths$asian`; and Native American or Alaskan Native = `r ftnotes$deaths$indian`.</b><br><br>

#### Ethnicity
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$deaths$ethn
```
<b>Note: Due to delays in reporting recent data are incomplete. Deaths with unknown ethnicity are not reported in the graph above (n=`r ftnotes$deaths$unk_ethn`).</b><br><br>

### Vaccination
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$deaths$vacc
```
<b>COVID-19 vaccines became available January 2021. `r ftnotes$deaths$ttl_after2021` (`r ftnotes$deaths$pct_after2021`) of the total deaths occurred after January 1st, 2021. In order to determine vaccination status, medical records and the New York State Immunization Information System are accessed. 69.0% (n=365) of the deaths that occurred in 2021 have a known vaccination status and 31.0% (n=164) have an unknown vaccination status.</b><br><br>

### Municipalities
```{r echo=FALSE, fig.height=5, message=FALSE, warning=FALSE, out.width='100%'}
plots$deaths$top_zip
```
<b>Note: Data is incomplete due to ongoing case investigation. Twenty most reported zip codes represented in graph (n=`r ftnotes$deaths$top_zip`) representing `r ftnotes$pct_deaths$pct_zip` of cases.</b><br><br>